Predicting Mental Workload: An Intermediate Fusion Deep Learning Approach

### 2021-04-20 manuscript/manuscript.tex

%Required packages
\usepackage[utf8]{inputenc}
\usepackage{apacite}
\usepackage{caption} 
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[textwidth=160mm, textheight=210mm, hmarginratio=1:1]{geometry}
\usepackage{geometry}
\usepackage[capposition=top]{floatrow}

@@

@@

%Title page 
\begin{titlepage}
\begin{center}
\LARGE{\textbf{The Realtime Assessment of Mental Workload by Means of Multiple Bio-Signals}}\\
\vspace*{2\baselineskip}
\Large{\textbf{Masterthesis Report}}\\
Methodology and Statistics for the Behavioral, Biomedical and Social Sciences\\
\vspace*{1\baselineskip}
Utrecht University\\
\vspace*{4\baselineskip}
{Bart-Jan Boverhof, 6000142}\\
\vspace*{1\baselineskip}
{\textbf{Thesis Supervisor}}\\
Prof.dr.ir. B.P. Veldkamp\\
\vspace*{1\baselineskip}
{\textbf{Date}}\\
January 7, 2021\\
\vspace*{1\baselineskip}
\end{center}
\end{titlepage}

%Introduction section
\section{Introduction}
The topic of mental workload is a widely studied phenomenon across a variety of different fields, amongst others the field of ergonomics \cite{young2015state}, human factors \cite{pretorius2007development} and cognitive neurosciences \cite{shuggi2017mental}. A commonly utilized definition of mental workload, hereafter referred to as simply "workload", is the demand placed upon one whilst one carries out a particular task. As rightfully pointed out by \citeA{de1996measurement}, the aforementioned definition is lacking, for it defines workload solely as a phenomenon external to the individual. Workload, however, requires to be recognized as a person-specific construct, for the amount of perceived workload ushered by a given task may vary substantially across individuals. \cite{de1996measurement}.

@@

With the current research endeavor we explore the feasibility of an approach to workload assessment by means of multiple modalities simultaneously, hereafter referred to as "multi-modular approach". We do so by combining data-streams stemming from three different modalities, being electroencephalography, photoplethysmography and galvanic skin response. Electroencephalography,  hereafter referred to as "EEG", is often utilized in the assessment of workload \cite{craik2019deep} \cite{berka2005evaluation} and found to be amongst the most adequately performing techniques within the field \cite{hogervorst2014combining}. Galvanic skin response, hereafter referred to as "GSR", equally so is a widely adopted approach towards workload assessment \cite{nourbakhsh2012using} \cite{zhou2015dynamic}. Lastly, heart-rate is a widely recognized indicator of workload, mostly obtained through photoplethysmography, hereafter referred to as "PPG" \cite{zhang2018evaluating} \cite{jimenez2018using}.

A deep learning approach towards modeling is adopted. A range of four different deep neural networks, hereafter referred to as "DNN's", are constructed and contrasted in performance: three of which are single-modular DNN's, i.e. networks each utilizing data from only a single modality. Thus,  one DNN solely utilizes the EEG modality,  one solely utilizes GSR and one solely utilizes PPG. The fourth network is a multi-modular network combining all three modalities into one. With this research we firstly sought to explore which modalities constitute the most adequate predictors of workload in an experimental setting, and secondly whether a multi-modular combination is preferable over the three simpler (but much less computationally demanding) single-modular DNN's. This assessment will be propelled by contrasting network performance, however when drawing conclusions, considerations regarding computational costs will be taken into consideration as well. All networks were constructed in Python, by utilizing the deep-learning toolbox PyTorch \cite{paszke2017automatic}.

@@

\subsection{Related Work} 
The following section provides an overview of preceding research regarding the assessment of workload by means of physiological bio-signals, with a focus on deep learning approaches. Attention is predominantly placed upon the most feasible network architectures, in addition to a range of model optimization techniques and to the data fusion strategy for the multi-modular DNN. The final part of this section will be dedicated to hyperparameter optimization.

@@

Research by \citeA{schirrmeister2017deep} contrasted the performance of several ConvNets against the widely utilized baseline method for EEG classification, filter bank common spatial pattern, hereafter referred to as "FBCSP". A deep ConvNet, a shallow ConvNet, a deep-shallow hybrid ConvNet and a residual ConvNet were contrasted with an FBCSP. Both the deep and shallow ConvNets were found to reach at least similar, and in some regards better classification results as compared with the FBCSP baseline approach. Altogether, a deep ConvNet with four convolutional-max-pooling blocks was found to perform best, exhibiting an accuracy of 92.4 \% \cite{schirrmeister2017deep}.

@@

\subsubsection{Photoplethysmography (PPG)}
Research by \citeA{biswas2019cornet} explored a deep learning approach towards PPG classification, with the objective to perform both bio-metric identification and obtain heart rate information. Exceptional results were realized with a DNN, attaining an average accuracy of 96 \% \cite{biswas2019cornet}. This performance was realized with a ConvNet-LSTM hybrid, incorporating two convolutional max-pooling blocks followed by two LSTM layers.

\subsubsection{Multi-modular Fusion Strategy}  
When conducting a multi-modular deep learning approach,  information streams stemming from the different modalities are required to be combined, i.e. "fused", at a certain point in the network in order to ultimately result in a single prediction of workload. Fusion can be done conforming several strategies. Three strategies as proposed by \citeA{ramachandram2017deep} have been considered.

@@

Lastly, intermediate fusion is the most widely employed fusion strategy for multi-modular deep-learning problems. Data streams are usually fused by a concatenation layer, joining the outputs of the separately defined network parts of each modality. This results in a single joint deep neural network. Several "higher-order layers" are usually defined in between the concatenation layer and the ultimate classification. The depth of the fusion, i.e. the specified number of higher-order layers, can be chosen conform to the specific circumstances, posing intermediate fusion to be the most flexible and therefore the most widely adopted fusion strategy \cite{ramachandram2017deep}.

Indeed, when consulting the literature, one is prone to conclude that intermediate fusion strategies are the most prevailing for multi-modular deep learning approaches. When taking an intermediate fusion approach, the higher-order part of the network needs to be designed and established, for which several previous research endeavors are considered.  \citeA{rastgoo2019automatic} utilized a multi-modular ConvNet approach, and fused the modalities by concatenation, followed with two LTSM layers and two dense layers. A simpler approach is adopted by \citeA{han2020classification}, who utilized an intermediate fusion approach solely consisting of several fully connected dense layers. Lastly, \citeA{dolmans2020perceived} took a relatively deep intermediate fusion approach, consisting of two dense layers, two convolutional layers followed by another two dense layers.

@@

\subsubsection{Single-modular Network}
All single-modular network architectures took inspiration from previously delineated research \cite{schirrmeister2017deep} \cite{dolmans2020perceived} \cite{sun2019hybrid} \cite{biswas2019cornet}. In contrast with most of these previous endeavors, we opted for a ConvNet-only approach, i.e. without LSTM-layers, for the objective of the current research is not of a time-series-forecasting-kind. An experimental scenario wherein the course of actions over the duration of the experiment always unfold in a similar manner are inclined to benefit from acknowledging this time-series-like nature, hence benefiting from LSTM. An example of such a scenario would be a pilot in flight: whilst workload arousing phenomena that occur during a flight could differ across takes, the general chronological structure of the scenario always adheres to a similar blueprint. I.e., starting with ascend, likely to arouse a peak in workload, and closing with touch down, equally so likely to arouse a peak in workload. Our experimental scenario doesn't adhere to such a set in stone unfolding, for the order in which the scenario's with variable difficulty were presented was entirely randomized.  Hence, for there is no chronologically consistent development of the experimental scenario, we opted for a ConvNet approach without opting to learn from the development over time, thus without LSTM-layers.

The three single-modular DNN architectures are depicted alongside one another as Figure \ref{fig:singlearchitecture}. Each of the single-modular networks can be distinguished by two separate part, being the convolutional part and the prediction part.  The objective of the convolutional part is to extract features from the data. For each network, this part consists of four convolutional blocks, each compromising a convolutional layer, a batch normalization layer, the Exponential Linear Unit, hereafter referred to as "ELU", activation function, and a max-pooling layer respectively. The prediction part compromises a flatten-layer,  with the objective of representing all input into one dimensional shape, followed by a dropout layer, a dense layer, a Rectified linear activation, hereafter referred to as "RELU", closed by another dense layer. An overview of the amount of utilized filters / neurons for each layer of each network is provided in Table \ref{table:modelvariations}.

Consistent with training, HPO was equally so done for each respondent individually, resulting in 27 optimized sets of parameters for the EEG networks, 27 for the PPG networks and 27 for the GSR networks.We optimized the learning rate, dropout rate and the amount of neurons for the first dense layer in 50 trials per network.  For training, a total of 600 epochs have been made through the data for each network independently. All training and HPO of the single-modular networks has been done on a V100 GPU,  offered by the GPU cloud service Google Collab.  Total running time for the single-modular networks was about four days, the most of which was absorbed by HPO.

\newgeometry{margin=2cm}
\begin{figure}
\caption{The Three Single-modular Network Architectures}
\bigskip
\includegraphics[scale=0.7]{single_model_architecture}
\label{fig:singlearchitecture}
\floatfoot{\emph{Note:} Each network constitutes a convolutional part of four convolutional blocks, and a classification part of two fully connected dense layers. }
\end{figure}
\restoregeometry

\newgeometry{margin=1.6cm}
\bgroup
\def\arraystretch{1.6}%  
\begin{table}[h]
\caption{Network Sizes}
\label{table:modelvariations}
\begin{tabular}{lllll}
\hline
        & EEG-Net/Part  & GSR-Net/Part  & PPG-Net/Part  & Multi-Modular Head-Net     \\ \hline
 Single-Modular-Nets & Dense Out: 1 & Dense Out: 1 & Dense Out: 1 &   \\
        & Dense 1: \emph{hpo} & Dense 1: \emph{hpo} & Dense 1: \emph{hpo} &   \\
        \vspace{1.5ex}        
        & Drop out: \emph{hpo} & Drop out: \emph{hpo} & Drop out: \emph{hpo} &   \\
        & Conv4: 200 & Conv4: 128 & Conv4: 128 &    \\ 
        & Conv3: 100 & Conv3: 64 & Conv3: 64 &    \\ 
        & Conv2 50 & Conv2: 32 & Conv2: 32 &    \\ 
        \vspace{3ex}
        & Conv1: 25 & Conv1: 16 & Conv1: 16 &  \\         
  Multi-Modular-Net & Dense 1: \emph{input} & Dense 1: \emph{input} & Dense 1: \emph{input} & Dense Out: 1  \\
        & Conv4: 200 & Conv4: 128 & Conv4: 128  &  Dense 3: \emph{hpo} \\ 
        & Conv3: 100 & Conv3: 64 & Conv3: 64 & Dense 2: \emph{hpo}  \\ 
        & Conv2 50 & Conv2: 32 & Conv2: 32 & Drop out: \emph{hpo}    \\ 
        \vspace{3ex}
        & Conv1: 25 & Conv1: 16 & Conv1: 16 &    \\   \hline
\end{tabular}
\vspace{2ex}
\begin{doublespacing}

@@

\subsubsection{Multi-modular Network}
The network architecture utilized for the multi-modular approach was determined by combining the previously characterized single-modular networks with insights drawn from previous research, in particular \cite{han2020classification}. A visual representation of the multi-modular network is depicted as Figure \ref{fig:multiarchitecture}. An overview of the amount of utilized filters and neurons for each layer of the multi-modular network is provided in Table \ref{table:modelvariations}.

The network architecture for the single-modular parts of the network are separately defined entities of the multi-modular network, but are highly similar as compared with the previously delineated single-modular networks. An intermediate fusion strategy is adopted due to its highly flexible nature as compared with alternative fusion strategies. The output of all three distinct parts in the network were flattened, followed by a dense layer and RELU activation.  Flattening was nescessary such that all inputs were represented in one-dimensional space, after which concatenation was possible. The prediction part, also referred to as "Head-Network", consists of three dense layers, alternated with dropout, batch normalization and RELU activation.

Again, consistent with training, HPO was equally so done for each respondent individually, resulting in 27 optimized sets of parameters each of the 27 multi-modular networks. We optimized the learning rate, dropout rate and the amount of neurons for the first and second dense layers in 18 trials per network.  For training, a total of 250 epochs have been made through the data for each network independently. All training and HPO has been done on a Cloud TPU v2,  offered by the GPU cloud service Google Collab Pro.  Total running time for the multi-modular networks was about 12 days, the most of which absorbed by HPO.

\newgeometry{margin=2cm}
\begin{figure}
\caption{The Multi-modular Network Architecture}
\bigskip
\includegraphics[scale=0.7]{multi_model_architecture}
\label{fig:multiarchitecture}
\floatfoot{\emph{Note:} The network constitutes a convolutional part of four convolutional blocks for each of distinct parts of the networks. All three distinct parts were flatenned, after which concatenation was possible.  The prediction part of the network constitutes several dense layers.}
\end{figure}
\restoregeometry

@@

Mean Absolute Error, hereafter referred to as "MAE",  of a DNN is the first utilized performance metric, defined as Equation \ref{eq:mae},

@@

Root Mean Squared Error, hereafter referred to as RMSE, is the second utilized performance metric, defined as Equation \ref{eq:rmse},

@@

Finally, the Pearson Correlation Coefficient, hereafter referred to as simply "correlation", constitutes the third utilized performance metric defined as Equation  \ref{eq:corr},

\begin{equation}
\label{eq:corr}
\rho_{X,Y} = \frac{cov(X,Y)}{\sigma_{X} \sigma_{Y}}
\end{equation}
where $x$ refers to the predicted values for the test windows,  $y$ to the true labels of test windows and $\sigma_{X}\, \& \,\sigma_{Y}$ refer to the standard deviation of $x$ and $y$ respectively.  Conceptually, correlation indicates the coherence between predictions and labels, hereby posing an alternative model performance metric.

\newpage
\section{Results}
Before illustrating results,  a short elaboration on the testing procedure will be provided.  Windows of eight seconds each were cut, resulting in a total of 9474 windows over all respondents, i.e. on average about 351 windows per respondent. For each network for each respondent individually,  80\% of all windows have been used for training, 10\% for validation within the training process and 10\% for assessing performance of the trained DNN's. This results in an average of 35 windows for assessing DNN performance per respondent, i.e. a total of 969 windows in total.  Partitioning windows into train-, validation- and testing -sets was done systematically by selecting every \textit{n-th} window, such that windows over the entire duration of the experiment were represented equally in each partition.

The distribution of all 9474 window labels, aggregated for all respondents,  is depicted as Figure \ref{fig:labels}. It becomes readily apparent that distribution of the window labels displays a noticeably right-skewed tendency, with a median of 5.75 and a 12.75 90\% upper-bound, implying that higher workload windows are relatively uncommon within both train and test partitions.  
\vspace{0.2cm}

\begin{figure}[h]
\caption{Distribution Window Labels}
\includegraphics[scale=0.42]{labels.png}
\label{fig:labels}
\floatfoot{\emph{Note:} The solid line represents the median, whereas the two dashed line represent the 10\% and 90\% quantiles respectively.}
\end{figure}

\subsection{Single-modular Network Performance}

\newgeometry{margin=1.4cm}
\begin{figure}
\caption{Mean Absolute Error (MAE) per modality}
\bigskip
\includegraphics[scale=0.7]{mae_hist.png}
\label{fig:mae_hist}
\floatfoot{\emph{Note:} The solid line represents the medians, whereas the two dashed line represent the 10\% and 90\% quantiles respectively.}
\end{figure}
\restoregeometry

\newgeometry{margin=1.1cm}
\begin{figure}
\caption{Mean Absolute Error (MAE) against label size}
\bigskip
\includegraphics[scale=0.67]{mae_scatter.png}
\label{fig:mae_scatter}
\end{figure}
\restoregeometry

\subsection{Multi-modular Network Performance}

Results EEG networks 
\begin{itemize}
\item Correlation predictions and labels: 0.68788.
\item Mean asbsolute error: 0.11029.
\item Mean absolute error on the original scale: 2.20586.
\end{itemize}
\bigskip

Results PPG networks 
\begin{itemize}
\item Correlation predictions and labels: 0.52998.
\item Mean asbsolute error: 0.13587.
\item Mean absolute error on the original scale: 2.71747.
\end{itemize}
\bigskip

Results GSR networks 
\begin{itemize}
\item Correlation predictions and labels: 0.65339.
\item Mean asbsolute error: 0.11883.
\item Mean absolute error on the original scale: 2.37665.
\end{itemize}
\bigskip

Results Multi-Modality networks 
\begin{itemize}
\item Correlation predictions and labels: 0.60864.
\item Mean asbsolute error: 0.12848.
\item Mean absolute error on the original scale: 2.56964.
\end{itemize}

\section{Limitations}

\begin{itemize}
  \item  More HPO could have been done, to get even better performing models but we had limited computatioal power.   
  \item This approach costs a lot of computational power to both train and do HPO. Especially since person-specific models are fitted. This upholds mostly for the multi-modal network.
  \item Due to the sheer size of the trained model and the fact that models are trained person specific, storage sizes of the models are huge. This could be unpractical.  This upholds mostly for the multi-modal network.
  \item Possibly different window sizes per modality could lead to better results.  * Insert some of the sources here that suggest different window sizes per modality* . This was beyond the scope of this project.
   \item How generizable is this to other situations for the same person (since things such as PPG are context dependent (like for example regarding how one ate).
   \item Long runtimes,  even when testing. This upholds only for the multi network. The reason is likely the sheer size of these networks, which for some persons are blown up due to device lagging. 
   \item hardware failure?
   \item The fact that we use TLX as labels.
\end{itemize}

>>>>

%Required packages
\usepackage[utf8]{inputenc}
\usepackage{apacite}
\usepackage{caption} 
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[textwidth=160mm, textheight=210mm, hmarginratio=1:1]{geometry}
\usepackage{geometry}
\usepackage[capposition=top]{floatrow}
\usepackage{authblk}

@@

%Frontpage specifications 
\title{Predicting  Mental Workload: An Intermediate Fusion Deep Learning Approach}

\author[1]{Bart-Jan Boverhof \thanks{b.boverhof@students.uu.nl}}
\author[2]{Bernard P. Veldkamp\thanks{b.p.veldkamp@utwente.nl}}
\affil[1]{\normalsize Faculty of Behavioral, Management and Social Sciences, University of Twente}
\affil[2]{\normalsize Faculty of Social and Behavioural Sciences, Utrecht University}

\renewcommand\Authands{ and }
\date{\today}

@@

\maketitle

\begin{abstract}
Morbi tempor congue porta. Proin semper, leo vitae faucibus dictum, metus mauris lacinia lorem, ac congue leo felis eu turpis. Sed nec nunc pellentesque, gravida eros at, porttitor ipsum. Praesent consequat urna a lacus lobortis ultrices eget ac metus. In tempus hendrerit rhoncus. Mauris dignissim turpis id sollicitudin lacinia. Praesent libero tellus, fringilla nec ullamcorper at, ultrices id nulla. Phasellus placerat a tellus a malesuada.
\end{abstract}

\hspace*{3.6mm}\textit{Keywords:} lorem, ipsum, dolor, sit amet, lectus % Keywords

\newpage
\section{Introduction}
The topic of mental workload is a widely studied phenomenon across a variety of different fields, amongst others the field of ergonomics \cite{young2015state}, human factors \cite{pretorius2007development} and cognitive neurosciences \cite{shuggi2017mental}. A commonly utilized definition of mental workload, hereafter referred to as simply "workload", is the demand placed upon one whilst one carries out a particular task. As rightfully pointed out by \citeA{de1996measurement}, the aforementioned definition is lacking, for it defines workload solely as a phenomenon external to the individual. Workload, however, requires to be recognized as a person-specific construct, for the amount of perceived workload ushered by a given task may vary substantially across individuals. \cite{de1996measurement}.

@@

With the current research endeavor we explore the feasibility of an approach to workload assessment by means of multiple modalities simultaneously, hereafter referred to as "multimodal approach". We do so by combining data-streams stemming from three different modalities, being electroencephalography, photoplethysmography and galvanic skin response. Electroencephalography,  hereafter referred to as "EEG", is often utilized in the assessment of workload \cite{craik2019deep} \cite{berka2005evaluation} and found to be amongst the most adequately performing techniques within the field \cite{hogervorst2014combining}. Galvanic skin response, hereafter referred to as "GSR", equally so is a widely adopted approach towards workload assessment \cite{nourbakhsh2012using} \cite{zhou2015dynamic}. Lastly, heart-rate is a widely recognized indicator of workload, mostly obtained through photoplethysmography, hereafter referred to as "PPG" \cite{zhang2018evaluating} \cite{jimenez2018using}.

A deep learning approach towards modeling is adopted. A range of four different deep neural networks, hereafter referred to as "DNN's", are constructed and contrasted in performance: three of which are unimodal DNN's, i.e. networks each utilizing data from only a single modality. Thus,  one DNN solely utilizes the EEG modality,  one solely utilizes GSR and one solely utilizes PPG. The fourth network is a multimodal network combining all three modalities into one. With this research we firstly sought to explore which modalities constitute the most adequate predictors of workload in an experimental setting, and secondly whether a multimodal combination is preferable over the three simpler (but much less computationally demanding) unimodal DNN's. This assessment will be propelled by contrasting network performance, however when drawing conclusions, considerations regarding computational costs will be taken into consideration as well. All networks were constructed in Python, by utilizing the deep-learning toolbox PyTorch \cite{paszke2017automatic}.

@@

\subsection{Related Work} 
The following section provides an overview of preceding research regarding the assessment of workload by means of physiological bio-signals, with a focus on deep learning approaches. Attention is predominantly placed upon the most feasible network architectures, in addition to a range of model optimization techniques and to the data fusion strategy for the multimodal DNN. The final part of this section will be dedicated to hyperparameter optimization.

@@

Research by \citeA{schirrmeister2017deep} contrasted the performance of several ConvNets against the widely utilized baseline method for EEG classification, filter bank common spatial pattern, hereafter referred to as "FBCSP". A deep ConvNet, a shallow ConvNet, a deep-shallow hybrid ConvNet and a residual ConvNet were contrasted with an FBCSP. Both the deep and shallow ConvNets were found to reach at least similar, and in some regards better classification results as compared with the FBCSP baseline approach. Altogether, a deep ConvNet with four convolutional-max-pooling blocks was found to perform best, exhibiting an accuracy of 92.4\% \cite{schirrmeister2017deep}.

@@

\subsubsection{Photoplethysmography (PPG)}
Research by \citeA{biswas2019cornet} explored a deep learning approach towards PPG classification, with the objective to perform both bio-metric identification and obtain heart rate information. Exceptional results were realized with a DNN, attaining an average accuracy of 96\% \cite{biswas2019cornet}. This performance was realized with a ConvNet-LSTM hybrid, incorporating two convolutional max-pooling blocks followed by two LSTM layers.

\subsubsection{Multimodal Fusion Strategy}  
When conducting a multimodal deep learning approach,  information streams stemming from the different modalities are required to be combined, i.e. "fused", at a certain point in the network in order to ultimately result in a single prediction of workload. Fusion can be done conforming several strategies. Three strategies as proposed by \citeA{ramachandram2017deep} have been considered.

@@

Lastly, intermediate fusion is the most widely employed fusion strategy for multimodal deep-learning problems. Data streams are usually fused by a concatenation layer, joining the outputs of the separately defined network parts of each modality. This results in a single joint deep neural network. Several "higher-order layers" are usually defined in between the concatenation layer and the ultimate classification. The depth of the fusion, i.e. the specified number of higher-order layers, can be chosen conform to the specific circumstances, posing intermediate fusion to be the most flexible and therefore the most widely adopted fusion strategy \cite{ramachandram2017deep}.

Indeed, when consulting the literature, one is prone to conclude that intermediate fusion strategies are the most prevailing for multimodal deep learning approaches. When taking an intermediate fusion approach, the higher-order part of the network needs to be designed and established, for which several previous research endeavors are considered.  \citeA{rastgoo2019automatic} utilized a multimodal ConvNet approach, and fused the modalities by concatenation, followed with two LTSM layers and two dense layers. A simpler approach is adopted by \citeA{han2020classification}, who utilized an intermediate fusion approach solely consisting of several fully connected dense layers. Lastly, \citeA{dolmans2020perceived} took a relatively deep intermediate fusion approach, consisting of two dense layers, two convolutional layers followed by another two dense layers.

@@

\subsubsection{Unimodal Network}
All unimodal network architectures took inspiration from previously delineated research \cite{schirrmeister2017deep} \cite{dolmans2020perceived} \cite{sun2019hybrid} \cite{biswas2019cornet}. In contrast with most of these previous endeavors, we opted for a ConvNet-only approach, i.e. without LSTM-layers, for the objective of the current research is not of a time-series-forecasting-kind. An experimental scenario wherein the course of actions over the duration of the experiment always unfold in a similar manner are inclined to benefit from acknowledging this time-series-like nature, hence benefiting from LSTM. An example of such a scenario would be a pilot in flight: whilst workload arousing phenomena that occur during a flight could differ across takes, the general chronological structure of the scenario always adheres to a similar blueprint. I.e., starting with ascend, likely to arouse a peak in workload, and closing with touch down, equally so likely to arouse a peak in workload. Our experimental scenario doesn't adhere to such a set in stone unfolding, for the order in which the scenario's with variable difficulty were presented was entirely randomized.  Hence, for there is no chronologically consistent development of the experimental scenario, we opted for a ConvNet approach without opting to learn from the development over time, thus without LSTM-layers.

The three unimodal DNN architectures are depicted alongside one another as Figure \ref{fig:singlearchitecture}. Each of the unimodal networks can be distinguished by two separate part, being the convolutional part and the prediction part.  The objective of the convolutional part is to extract features from the data. For each network, this part consists of four convolutional blocks, each compromising a convolutional layer, a batch normalization layer, the Exponential Linear Unit, hereafter referred to as "ELU", activation function, and a max-pooling layer respectively. The prediction part compromises a flatten-layer,  with the objective of representing all input into one dimensional shape, followed by a dropout layer, a dense layer, a Rectified linear activation, hereafter referred to as "RELU", closed by another dense layer. An overview of the amount of utilized filters / neurons for each layer of each network is provided in Table \ref{table:modelvariations}.

Consistent with training, HPO was equally so done for each respondent individually, resulting in 27 optimized sets of parameters for the EEG networks, 27 for the PPG networks and 27 for the GSR networks.We optimized the learning rate, dropout rate and the amount of neurons for the first dense layer in 50 trials per network.  For training, a total of 600 epochs have been made through the data for each network independently. All training and HPO of the unimodal networks has been done on a V100 GPU,  offered by the GPU cloud service Google Collab.  Total running time for the unimodal networks was about four days, the most of which was absorbed by HPO.

\newgeometry{margin=2cm}
\begin{figure}
\caption{The Three Unimodal Network Architectures}
\bigskip
\includegraphics[scale=0.69]{single_model_architecture}
\label{fig:singlearchitecture}
\floatfoot{\emph{Note:} Each network constitutes a convolutional part of four convolutional blocks, and a classification part of two fully connected dense layers. }
\end{figure}
\restoregeometry

\newgeometry{margin=1.6cm}
\bgroup
\def\arraystretch{1.6}%  
\begin{table}[h]
\caption{Network Sizes}
\label{table:modelvariations}
\begin{tabular}{lllll}
\hline
        & EEG-Net/Part  & GSR-Net/Part  & PPG-Net/Part  & Multimodal Head-Net     \\ \hline
 Unimodal-Nets & Dense Out: 1 & Dense Out: 1 & Dense Out: 1 &   \\
        & Dense 1: \emph{hpo} & Dense 1: \emph{hpo} & Dense 1: \emph{hpo} &   \\
        \vspace{1.5ex}        
        & Drop out: \emph{hpo} & Drop out: \emph{hpo} & Drop out: \emph{hpo} &   \\
        & Conv4: 200 & Conv4: 128 & Conv4: 128 &    \\ 
        & Conv3: 100 & Conv3: 64 & Conv3: 64 &    \\ 
        & Conv2 50 & Conv2: 32 & Conv2: 32 &    \\ 
        \vspace{3ex}
        & Conv1: 25 & Conv1: 16 & Conv1: 16 &  \\         
  Multimodal-Net & Dense 1: \emph{input} & Dense 1: \emph{input} & Dense 1: \emph{input} & Dense Out: 1  \\
        & Conv4: 200 & Conv4: 128 & Conv4: 128  &  Dense 3: \emph{hpo} \\ 
        & Conv3: 100 & Conv3: 64 & Conv3: 64 & Dense 2: \emph{hpo}  \\ 
        & Conv2 50 & Conv2: 32 & Conv2: 32 & Drop out: \emph{hpo}    \\ 
        \vspace{3ex}
        & Conv1: 25 & Conv1: 16 & Conv1: 16 &    \\   \hline
\end{tabular}
\vspace{2ex}
\begin{doublespacing}

@@

\subsubsection{Multimodal Network} \label{section:multimodal}
The network architecture utilized for the multimodal approach was determined by combining the previously characterized unimodal networks with insights drawn from previous research, in particular \cite{han2020classification}. A visual representation of the multimodal network is depicted as Figure \ref{fig:multiarchitecture}. An overview of the amount of utilized filters and neurons for each layer of the multimodal network is provided in Table \ref{table:modelvariations}.

The network architecture for the unimodal parts of the network are separately defined entities of the multimodal network, but are highly similar as compared with the previously delineated unimodal networks. An intermediate fusion strategy is adopted due to its highly flexible nature as compared with alternative fusion strategies. The output of all three distinct parts in the network were flattened, followed by a dense layer and RELU activation.  Flattening was necessary such that all inputs were represented in one-dimensional space, after which concatenation was possible. The prediction part, also referred to as "Head-Network", consists of three dense layers, alternated with dropout, batch normalization and RELU activation.

Again, consistent with training, HPO was equally so done for each respondent individually, resulting in 27 optimized sets of parameters each of the 27 multimodal networks. We optimized the learning rate, dropout rate and the amount of neurons for the first and second dense layers in 18 trials per network.  For training, a total of 250 epochs have been made through the data for each network independently. All training and HPO has been done on a Cloud TPU v2,  offered by the GPU cloud service Google Collab Pro.  Total running time for the multimodal networks was about 12 days, the most of which absorbed by HPO.

\newgeometry{margin=2cm}
\begin{figure}
\caption{The Multimodal Network Architecture}
\bigskip
\includegraphics[scale=0.69]{multi_model_architecture}
\label{fig:multiarchitecture}
\floatfoot{\emph{Note:} The network constitutes a convolutional part of four convolutional blocks for each of distinct parts of the networks. All three distinct parts were flatenned, after which concatenation was possible.  The prediction part of the network constitutes several dense layers.}
\end{figure}
\restoregeometry

@@

The Mean Absolute Error, hereafter referred to as "MAE",  of a DNN is the first utilized performance metric, defined as Equation \ref{eq:mae}

@@

The Root Mean Squared Error, hereafter referred to as RMSE, is the second utilized performance metric, defined as Equation \ref{eq:rmse}

@@

Finally, the Pearson Correlation Coefficient, hereafter referred to as simply "correlation", constitutes the third utilized performance metric defined as Equation  \ref{eq:corr}

\begin{equation}
\label{eq:corr}
r_{X,Y} = \frac{cov(X,Y)}{\sigma_{X} \sigma_{Y}}
\end{equation}
where $x$ refers to the predicted values for the test windows,  $y$ to the true labels of test windows and $\sigma_{X}\, \& \,\sigma_{Y}$ refer to the standard deviation of $x$ and $y$ respectively.  Conceptually, correlation indicates the coherence between predictions and labels, hereby posing an alternative model performance metric.

\newpage
\section{Results}
Before illustrating DNN results,  a short elaboration on the testing procedure will be conducted.  The data streams constituting the entire length of the experiment were cut into windows of eight seconds, resulting in an average of roughly 351 windows per respondent. For each network for each respondent individually,  80\% of all windows have been used for training, 10\% for validation within the training process and 10\% for the assessment of performance. This resulted in an average of 35 windows for assessing DNN performance per respondent.  Partitioning windows into train-, validation- and testing -sets was done systematically by selecting every \textit{n-th} window, such that windows over the entire duration of the experiment were represented equally in each partition.  For each respondent, the same partitions were used across all four DNN's.

The distribution of all 9474 window labels, aggregated for all respondents,  is depicted as Figure \ref{fig:labels}. It becomes readily apparent that distribution of the window labels displays a noticeably right-skewed tendency, with a median of 5.75 and an 90\% upper-bound quantatile of 12.75, implying that higher workload windows are relatively uncommon within all partitions.  
\vspace{0.1cm}

\begin{figure}[h]
\caption{Distribution Window Labels}
\includegraphics[scale=0.39]{labels.png}
\label{fig:labels}
\floatfoot{\emph{Note:} The solid line represents the median, whereas the two dashed line represent the 10\% and 90\% quantiles respectively.}
\end{figure}

\subsection{Deep Neural Network Performance}
Depicted in Table \ref{table:results} are the performance metrics for each of the four network architectures. The most conspicuous result that becomes apparent is that the multimodal architecture performs consistently worse on all metrics as compared with both the unimodal EEG and GSR networks.

\vspace{0.3cm}
\bgroup
\def\arraystretch{1.6}%  
\begin{table}[h]
\caption{Deep Neural Network Performance Metrics}
\label{table:results}
\begin{tabular}{lllll}
\hline
                                                                               & EEG\hspace{0.5cm} & PPG\hspace{0.5cm} & GSR\hspace{0.5cm} & Multi\hspace{0.5cm} \\ \hline
\begin{tabular}[c]{@{}l@{}}Mean Absolute Error\\ \hspace{1.2cm}(\textit{0-1 scale})\end{tabular} & 0.110 & 0.136 & 0.119 & 0.128  \\
\begin{tabular}[c]{@{}l@{}}Mean Absolute Error\\ \hspace{0.25cm}(\textit{original 0-20 scale})\end{tabular} & 2.206 & 2.717 & 2.377 & 2.570 \\
Root Mean Square Error                                                         & 0.154 & 0.180 & 0.159 & 0.167 \\
Pearson Correlation                                                                    & 0.688 & 0.530 & 0.653 & 0.609 \\ \hline
\end{tabular}
\end{table}

Performance in terms of MAE for the EEG architecture is highest of all other architectures, with $MAE = 0.110$ with a standard deviation, hereafter referred to as "\emph{sd}" of $\sigma = 0.154$.  The architecture demonstrating the next highest MAE is the GSR architecture, displaying an $MAE = 0.119$ with an sd of $\sigma = 0.159$.  The multimodal networks is observed to display an $MAE = 0.128$ with an sd of $\sigma = 0.167$.  Lastly, the PPG networks exhibits the highest error of $MAE = 0.136$ with an sd of $\sigma = 0.179$.  Results in terms of RMSE adhere to a similar pattern as MAE do, which is not surprising given that the RMSE and MAE are similar. The difference is that the RMSE values for all networks reside slightly higher, which is not surprising for the punishes high deviations more severely. The correlation between predictions and labels are indicative of a similar pattern as compared with the other metrics. However despite similarity, an interesting disparity is that the differences across the four architectures seems to be more substantial. The value for the EEG networks are observed to be $r=0.688$. Equally so, the GSR networks display a fairly strong correlation of $r=0.653$. Again, both the multimodal and GSR architectures are the least adequately performing in terms of correlation, displaying $r=0.609$ and $r=0.530$ respectively.

An independent samples t-test was conducted to formally test differences in network performance. To be precise, we statistically tested the difference in absolute error per window across several pairs of architectures (two-sided with $\alpha = 0.05$).  The absolute error for the unimodel EEG architecture was found to be significantly lower as compared with the multimodel architecture $t(1936)=-2.566, p=0.01$. We did not find a significant difference in absolute error between the GSR- and multimodal -architecture $t(1936)=-1.188, p=0.24$, and equally so no significant difference was found between the GSR- and multi- architectures $t(1936)=-1.463, p = 0.144$. Lastly, absolute error for the PPG architecture was found to be significantly lower as compared with the GSR architecture $t(1936)=-2.636 , p = 0.008$, but no difference was found between the GSR- and multimodal -architecture $t(1936)=-1.164 , p = 0.245$.

To explore the difference in performance into more detail, a graphical visualization of the prediction errors for each of the four network architectures is depicted as Figure \ref{fig:mae_hist}.
